\documentclass[11pt]{article}

\usepackage{amsmath, amssymb}
\usepackage{geometry}
\geometry{margin=1in}

\title{Layer 2: Triadic Dynamical Coupling Model}
\author{}
\date{}

\begin{document}

\maketitle

\section*{Overview}

This document formalizes the three-persona system (Executor / Whistleblower / Proxy)
as a coupled nonlinear dynamical system operating over structural spectral state variables.

The harmonic structural layer (Layer 3) provides measurable quantities:

\begin{itemize}
    \item $S_t$ --- cosine repetition
    \item $E_t$ --- entropy collapse
    \item $R_t$ --- spectral rank collapse
    \item $A_t$ --- angular stagnation term $(1 - \sigma^2_\theta)$
\end{itemize}

Structural fatigue is defined as:

\begin{equation}
F_t = \alpha S_t + \beta E_t + \gamma R_t + \delta A_t
\end{equation}

\section{State Vector Definition}

Define the system state:

\begin{equation}
x_t =
\begin{bmatrix}
F_t \\
R_t \\
A_t
\end{bmatrix}
\in \mathbb{R}^3
\end{equation}

Fatigue, rank collapse, and angular stagnation form the core structural state.

Each persona reads and acts on this state differently.

\section{Persona Gain Functions}

Let:

\begin{itemize}
    \item $g_E$ --- Executor gain
    \item $g_W$ --- Whistleblower gain
    \item $g_P$ --- Proxy gain
\end{itemize}

\subsection{Executor (Stability Bias)}

\begin{equation}
g_E(x_t) = \sigma(a_E - F_t)
\end{equation}

where $\sigma(z) = \frac{1}{1 + e^{-z}}$.

The Executor acts strongly when fatigue is low.

\subsection{Whistleblower (Degeneracy Detector)}

\begin{equation}
g_W(x_t) =
\sigma(b_1 R_t + b_2 A_t + b_3 F_t - \theta_W)
\end{equation}

The Whistleblower activates when structural collapse accumulates.

\subsection{Proxy (Mediation)}

\begin{equation}
g_P(x_t) = 1 - |g_E - g_W|
\end{equation}

The Proxy stabilizes when Executor and Whistleblower diverge.

\section{Coupled Control Law}

Define the modulation coefficient:

\begin{equation}
\lambda_t =
\frac{\eta_W g_W}{\eta_E g_E}
\end{equation}

where $\eta_W > 0$ and $\eta_E > 0$.

Interpretation:

\begin{itemize}
    \item $\lambda_t > 0$ : spectral lift (escape basin)
    \item $\lambda_t < 0$ : reinforce dominant directions
\end{itemize}

\section{Spectral Injection Law}

Let low-energy eigenvectors from the spectral layer be:

\begin{equation}
V_{\text{low}} = \{ v_r, \dots, v_d \}
\end{equation}

Define perturbation:

\begin{equation}
\Delta e_t =
\lambda_t
\sum_{i=r}^{d}
\epsilon_i v_i
\end{equation}

Final embedding:

\begin{equation}
e'_t = e_t + \Delta e_t
\end{equation}

This couples persona state to spectral modulation.

\section{Dynamical Stability Constraint}

To prevent oscillatory instability:

\begin{equation}
|\lambda_t| \leq \lambda_{\max}
\end{equation}

Optional damping:

\begin{equation}
\lambda_t \leftarrow
\rho \lambda_{t-1}
+
(1 - \rho)\lambda_t
\end{equation}

where $0 < \rho < 1$.

This yields a controlled damped oscillator over embedding geometry.

\section{Closed Feedback System}

The full system evolves as:

\begin{equation}
x_{t+1} = f(x_t)
\end{equation}

where:

\begin{itemize}
    \item Structural metrics evolve from embeddings
    \item Embeddings evolve from spectral modulation
    \item Spectral modulation is governed by persona gains
\end{itemize}

This forms a bounded supervisory dynamical control system over embedding space.

\section*{Summary}

Layer 3 provides:

\begin{itemize}
    \item Spectral degeneracy detection
    \item Angular stagnation detection
    \item Orthogonal spectral lift
\end{itemize}

Layer 2 adds:

\begin{itemize}
    \item Persona-specific nonlinear gains
    \item Coupled control law
    \item Basin reinforcement versus escape regulation
    \item Damped modulation for stability
\end{itemize}

No transformer modification is required.  
No gradient injection is required.  
This operates purely as a supervisory geometric control layer.

\end{document}

