\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, geometry, hyperref}
\geometry{margin=1in}

\title{Sovereign Execution Protocol (SEP-B\textsubscript{h}) for the Vessel}
\author{}
\date{}

\begin{document}
\maketitle

\section*{Introduction}

This document formalizes a minimal \textbf{Sovereign Execution Protocol (SEP)} for a personal Vessel system. The design prioritizes human sovereignty, temporal persistence, and controlled AI mode execution.

\section*{Persistent Direction}

We define a persistent trajectory vector, $\tau_{persistent}$, that anchors all activity and decisions of the Vessel:

\[
\tau_{dir} \equiv \tau_{persistent}
\]

\begin{itemize}
    \item \textbf{Persistent Direction Vector}: Carries continuity across sessions.
    \item Anchors all mode activity $(M_t, \Theta_t)$ to a single human-defined trajectory.
    \item Drift and deviation checks always reference $\tau_{persistent}$.
\end{itemize}

\section*{SEP-B\textsubscript{h} Definition}

\subsection*{1. State Representation}
\[
S_t = (M_t, \Theta_t, \tau_{persistent})
\]

\begin{itemize}
    \item $M_t$: Current AI mode.
    \item $\Theta_t$: Operational parameters at time $t$.
    \item $\tau_{persistent}$: Human-anchored trajectory vector.
\end{itemize}

\subsection*{2. Parameter Constraints}
\[
D_t = \| O_t - \tau_{persistent} \| \leq \kappa
\]

\begin{itemize}
    \item $O_t$: Observed operational vector at time $t$.
    \item $\kappa$: Maximum allowable deviation from persistent trajectory.
\end{itemize}

\subsection*{3. Mode Transition Rule}
\[
M_t \rightarrow M_{t+1} \implies \text{HumanInitiated only}
\]

\subsection*{4. Execution Check for Irreversible Operations}
\[
\text{Irreversible(Op)} \Rightarrow \text{Confirm(H)}
\]

\subsection*{5. Session Persistence}
\[
\tau_{persistent} \text{ stored and recalled automatically each session}
\]

\section*{Notes}

\begin{itemize}
    \item Ensures that the Vessel cannot self-derive long-term objectives beyond $\tau_{persistent}$.
    \item All optimization, entropic modulation, and pulse mechanisms must reference $\tau_{persistent}$ as the grounding vector.
    \item Human sovereignty is preserved explicitly across sessions.
\end{itemize}

\end{document}

